The \project{unRAVE} catalog

### 2016-09-08 article/unrave.tex
Finish conc

We have derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all \rave\ 
spectra, as well as detailed abundances for red giant branch stars, based on an
independent re-analysis of \rave\ spectra.  Our catalog provides an excellent
complement to the current \rave\ data release \citep{Kunder_2016}.  
When combined with the \tgas\ sample, these results may amount to a powerful
compendium for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
Inferences based on these results should recognize those caveats, and acknowledge that 
these results are subject to our explicit assumptions, some of which are provably
incorrect.

For practical purposes we adopted separate models: one for the giant branch and one
for the main-sequence.  A third model was used to derive relative weights for which
results to use.  The relative weighting we have used does not have any formal
interpretation as a likelihood or belief (in any sense): it was introduced for
practical reasons to identify systematic errors and combine results for multiple
models.  Because the relative weights have no formal interpretation, it is reasonable
to consider this method is as \emph{ad hoc} as any other approach.  Therefore, the
relative weighting has no warranty to be (formally) correct, and may introduce 
inconsistencies or systematic errors rather than
minimizing them.  Evidence of this is already present in our comparisons to
`gold standard' works \citep{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005},
where metal-poor dwarf stars appear as metal-poor sub-giants in our combined results,
but have reasonable labels (by any available statistic) as dwarfs in the main-sequence
model.  Clearly, the relative distance from the joint model to the giant model is 
slightly favoured (over the main-sequence model) for metal-poor dwarf stars. In this
sense, our approach towards combining labels from multiple models is sub-optimal.

If we only consider the results from individual models, there are a number of cautionary
remarks that stem from the construction of the training set.  The labels for red giant 
branch stars primarily come from \apogee, where previous successes with \thecannon\
have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}.
However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
a tapering-off in the test set --- where \emph{no} stars had reliably reported metallicities
below that of the training set --- which forced us to construct a heterogeneous training
set.  The metal-poor stars included in this sample are from high-resolution studies
\citep{Fulbright_2010,Ruchti_2011}, but it is unknown if the stellar parameters are
of high fidelity (e.g., we have limited quality metrics available), and there is no
guarantee that the stellar parameters \emph{or} abundances are on the same scale as
\apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.

If the metallicities of metal-poor giant stars are on the same scale as the \apogee\
abundances, there is a larger issue in verifying that the main-sequence metallicities
and giant branch metallicities are on the same scale.  The training set for the
main-sequence stars includes metallicities from a variety of sources --- including
\lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
no straightforward manner to ensure that these two models produce metallicities on
the same scale.  We see no systematic offset in metallicities of dwarf and giant
stars that overlap between \rave\ and the \ges\ survey, suggesting that if there is
a systematic offset, it must be small. Nevertheless, these are only verification checks based on $<1$\% of the
data, and there is currently insufficient data for us to prove both models are on
the same abundance scale.

For some of the most metal-poor giant stars in \rave\, we \emph{know} the abundances 
are not on the same scale as \apogee, because we were forced to adopt abundances for 
specific elements when they were unavailable.  In that sense, our assumption that the
training set contains accurate labels is clearly untrue (for \emph{at least} some stars).
Although we sought to adopt the mean
level of Galactic chemical enrichment at the overall metallicity, this is not a
representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
metallicity, there is no requirement for zero abundance spread.
More fundamentally, we are incorrectly assuming that the element \emph{must} be 
detectable in the photosphere of the star.  There may be no transition that is 
detectable in that star, even with zero noise, because it is too weak to have any 
effect on the spectrum.  In the most optimistic case, this can be considered to be
forcing the model to make use of correlated information between abundances.  In a
more representative (pessimistic) case, we are simply invoking what all abundances
should be at low metallicity.

@@

In Section~\ref{sec:method} we assumed that any fibre- or time-dependent variations
in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
\citet{Kordopatis_2013} note that the effective resolution of \rave\
spectra varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the 
effective resolution is a function of temperature variations, fibre-to-fibre
variations, and thus position on the CCD \citep{Steinmetz_2006}.
 For this reason we ought to expect our derived stellar
parameters or abundances to be correlated with the fibre number, or over
time.  If significant, the trend could produce systematically offset
stellar abundances solely due to the fibre number.  \citet{Kordopatis_2013}
conclude that resolution-based effects on the \rave\ stellar parameters
should be a second-order effect. We have not seen evidence of these 
resolution-based correlations in our results, however, we have only
performed cursory (non-exhaustive) experiments to investigate this issue.

\stub{assumptions that are probably wrong: fringing in high S/N spectra produces issues with the normalization and thus the spectrum. KCF to comment more on this}

The fibre fringing is another example of how test spectra that are
unusual from the training set can be projected as peculiar artefacts in the
label space.  In other words, unusual spectra can appear as `clumps' in
regions of parameter space that we could consider as being normal 
(e.g., solar-type stars).  We addressed this issue for the main-sequence
and giant models by using a joint model to calculate relative weights.
However, spectra that are unusual from the joint model training set can
still project as systematic artefacts.

Indeed, there are two known artefacts in our data that are relevant to this discusssion. The
first is a small over-density at the base of the giant branch, which is
likely a consequence of joining the 9-label and 3-label models.  The
second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
present in our training set, yet there are many in the \rave\ survey.
However, the \rave\ pre-processing pipeline \citep[\texttt{SPARV};][]{Steinmetz_2006,Zwitter_2008} performs
template matching against grids of cool \emph{and} hot stars, and therefore
we can use those results to identify hot stars. In Figure \ref{fig:hot-stars} 
we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
where each hexagonal bin is colored by the \emph{maximum} temperature
reported by \texttt{SPARV} for any star in that bin.  Hot stars dominate
at just above the turn-off point in our results.

\stub{this following paragraph needs rewording}
Based on a thorough interrogation of our results, we have adopted two boolean
columns to signify whether our labels meet specific quality constraints:
one marked \texttt{QC} to indicate reliable stellar parameters ($\teff$, $\logg$, [Fe/H]), and \texttt{QC\_ABUNDANCE} to indicate reliable chemical abundances.
For stellar atmospheric parameters to meet our quality constraints, we
require: $\chi_r^2 < 3$; $S/N >10$~pixel$^{-1}$; and \texttt{TeffS} $< 8000$~K
in order to remove hot stars that are outside the training set.  We
specified the quality constraint in chemical abundances to further
require that the iron abundance $[{\rm Fe/H}] > -1.5$, since this approximately
corresponds to the point where individual abundances will become undetectable
in \rave\ spectra, and the individual abundance labels will be driven by our
constraint to place them on the mean galactic chemical evolution trend.
These quality constraints should not be taken verbatim: judicious use of
the morphological classifications \citep{Matijevic_2012} or metadata from
the pre-processing pipelines is encouraged.

We have presented the largest collection of stellar abundances for stars in the 
first \gaia\ data release.  In total more than 1.69 million abundances are
available, with effective temperature $\teff$, surface gravity $\logg$ and 
metallicity available for \Nstarsqc\ stars.  Despite the caveats and limitations 
discussed here, our validation experiments and comparisons with high-resolution
spectroscopic studies suggests that our results have sufficient accuracy and
precision to be useful for exoplanet host stars studies 
\citep[c.f.,][]{Valenti_Fischer_2005}, investigations of the Milky Way disk 
\citep[c.f.,][]{Reddy_2003,Reddy_2006,Bensby_2014}, and a multitude of 
chemo-dynamic studies.  When combined with positions and 3D velocities from \gaia,
we are optimistic that the \unrave\ catalog will help advance understanding 
of the Milky Way's formation and evolution.

>>>>

We have performed an independent re-analysis of \Nspectra\ \rave\ spectra,
having derived atmospheric parameters ($\teff$, $\logg$, [Fe/H]) for all stars,
as well as detailed chemical abundances for red giant branch stars.  When 
combined with the \tgas\ sample, these results amount to a powerful compendium 
for chemo-dynamic studies of the Milky Way.  However, our analysis has caveats.
Inferences based on these results should recognize those caveats, and acknowledge 
that these results are subject to our explicit assumptions, some of which are provably
incorrect.

For practical purposes we adopted separate models: one for the giant branch and one
for the main-sequence.  A third model was used to derive relative weights for which
results to use.  The relative weighting we have used does not have any formal
interpretation as a likelihood or belief (in any sense): it was introduced for
practical reasons to identify systematic errors and combine results for multiple
models.  Because the relative weights have no formal interpretation, it is reasonable
to consider this method is as \emph{ad hoc} as any other approach.  The relative 
weighting has no warranty to be (formally) correct, and may introduce inconsistencies
or systematic errors rather than minimizing them.

If we only consider the results from individual models, there are a number of cautionary
remarks that stem from the construction of the training set.  The labels for red giant 
branch stars primarily come from \apogee, where previous successes with \thecannon\
have demonstrated that \apogee\ labels based on high S/N data can be of high fidelity
\citep{Ness_2015,Ness_2016,Ho_2016,Casey_2016b}.
However, the lack of metal-poor stars in the \apogee/\rave\ overlap sample produced
a tapering-off in the test set --- where \emph{no} stars had reliably reported metallicities
below that of the training set --- which forced us to construct a heterogeneous training
set.  The metal-poor stars included in this sample are from high-resolution studies
\citep{Fulbright_2010,Ruchti_2011}, but it is not known if the stellar parameters are
of high fidelity because we have limited quality metrics available.  Moreover, there
is no guarantee that the stellar parameters \emph{or} abundances are on the same scale
as \apogee\ \citep[and good reasons to believe they will not be; see][]{Smiljanic_2014}.

If the metallicities of metal-poor giant stars were on the same scale as the \apogee\
abundances, there is a larger issue in verifying that the main-sequence metallicities
and giant branch metallicities are on the same scale.  The training set for the
main-sequence stars includes metallicities from a variety of sources, including
\lamost, and the fourth \rave\ data release.  Even on expectation value, there is 
no straightforward manner to ensure that the main-sequence model and the red giant
branch model produce metallicities on the same scale.  We see no systematic offset
in metallicities of dwarf and giant stars that overlap between \rave\ and the \ges\
survey, suggesting that if there is a systematic offset, it must be small. 
Nevertheless, these are only verification checks based on $<1$\% of the data, and 
there is currently insufficient data for us to prove both models are on the same 
abundance scale.

For some of the most metal-poor giant stars in \rave\, we \emph{know} the abundances 
are not on the same scale as \apogee, because we were forced to adopt abundances for 
specific elements when they were unavailable.  Although we sought to adopt mean
level of Galactic chemical enrichment at a given overall metallicity, this is not a
representative abundance. Even if that is the \emph{mean} enrichment at that Galactic
metallicity, there is no requirement for zero abundance spread.  More fundamentally, 
we are incorrectly asserting that the element \emph{must} be detectable in the 
photosphere of the star.  There may be no transition that is detectable in that 
star, even with zero noise, because it is too weak to have any effect on the 
spectrum.  In the most optimistic case, this could be considered to be forcing the
model to make use of correlated information between abundances.  In a more 
representative (pessimistic) case, we are simply invoking what all abundances should 
be at low metallicity.

@@

In Section~\ref{sec:method} we assumed that any fibre- or time-dependent variations
in the \rave\ spectra are negligible.  This is provably incorrect. Indeed,
\citet{Kordopatis_2013} note that the effective resolution of \rave\ spectra
varies from $6{,}500 < \mathcal{R} < 8{,}500$, and that the effective 
resolution is a function of temperature variations, fibre-to-fibre variations,
and thus position on the CCD \citep{Steinmetz_2006}. For this reason we ought
to expect our derived stellar parameters or abundances to be correlated either
with the fibre number, with the observation date, or both.  If significant, 
the trend could produce systematically offset stellar abundances solely due to
the fibre used.  \citet{Kordopatis_2013} conclude that resolution-based effects 
on the \rave\ stellar parameters should be a second-order effect. We have not 
seen evidence of these resolution-based correlations in our results, however, 
we have only performed cursory (non-exhaustive) experiments to investigate 
this issue.

We have shown some potential outcomes when the test set spectra differ 
significantly from the spectra in the training set.  Test set spectra that is
`unusual' from the training set can be projected as peculiar artefacts in 
label space.  In other words, unusual spectra can appear as `clumps' in
regions of parameter space that we could consider as being normal (e.g.,
an over-density of solar-type stars).  We addressed this issue for the main-sequence
and giant models by using a third model (Section \ref{sec:a-simple-model}) to
calculate relative weights.  However, spectra that are unusual from the 
training set used in the simplest model still project as systematic artefacts.

Indeed, there are two known artefacts in our data that are relevant to this 
discussion. The first is a small over-density at the base of the giant branch, 
which is likely a consequence of joining the 9-label and 3-label models.  The
second has an astrophysical origin: there are no hot stars ($\teff > 8000$~K)
present in our training set, yet there are many in the \rave\ survey.
However, the \rave\ pre-processing pipeline 
\citep[\texttt{SPARV};][]{Steinmetz_2006,Zwitter_2008} performs template 
matching against grids of cool \emph{and} hot stars, and therefore we can use 
that information to identify hot stars.  In Figure \ref{fig:hot-stars} 
we show our derived effective temperatures $\teff$ and surface gravities $\logg$,
where each hexagonal bin is colored by the \emph{maximum} temperature
reported by \texttt{SPARV} for any star in that bin.  We show the maximum
temperature reported by \texttt{SPARV} to demonstrate that hot stars project
into a single clump in our label space --- near the turn-off --- in a region
where we may otherwise be deceived into thinking the observed over-density
is consistent with expectations from astrophysics.

We have included a boolean flag column in our catalog for quality control,
to signify whether a result is thought to be reasonable.  Specifically, 
stars where \texttt{SPARV} indicates a $\teff > 8000$~K do not meet our quality 
control flag.  For all other stars, we further require $\chi_r^2 < 3$ and
$S/N >10$~pixel$^{-1}$.  These quality constraints represent the minimum
acceptable constraints and should not be taken verbatim: judicious use of
the morphological classifications \citep{Matijevic_2012} or additional 
metadata from the \rave\ pre-processing pipelines is strongly encouraged.

We have presented the largest collection of stellar abundances for stars in the 
first \gaia\ data release.  In total we derive reliable stellar atmospheric 
parameters for \Nstarsqc, and report more than 1.69 million abundances.  Despite 
the caveats and limitations discussed here, our validation experiments and 
comparisons with high-resolution spectroscopic studies suggests that our results 
have sufficient accuracy and precision to be useful for chemo-dynamic studies 
that become imminently feasible only in the era of \gaia\ data.  Indeed, we
are optimistic that the \unrave\ catalog will advance understanding of the
Milky Way's formation and evolution.
